---
title: "Group Report Visualisation"
output:
  pdf_document: default
  html_document: default
---



```{r Load Packages, include=FALSE}
library(tidyverse)
library(reticulate)
library(lubridate)
library(reshape2)
library(glue)
```


# Pre-Processing (do not run this section as it has already been processed - proceed to Visualisation section)

```{python Reading Data in Python}
# Lets first make some columns - using pandas
import pandas as pd

# Import dataset into Python
df_tempTd_pd = pd.read_csv('../data/processed/data_used_to_build_model.csv')

# Converting DATETIME column to datetime64[ns]
df_tempTd_pd['DATETIME'] = pd.to_datetime(df_tempTd_pd['DATETIME'], format='%Y-%m-%d %H:%M:%S')

# Get info on data
df_tempTd_pd.info()
```

```{python Inspecting data}
# Inspect the data
df_tempTd_pd.head(5)
```


```{python Adding new features}
# Create new features
def create_new_features(df):
    df["HOUR"] = df["DATETIME"].dt.hour.astype('int8')
    df["ISO_DAYOFWEEK"] = df["DATETIME"].dt.isocalendar().day.astype('int8')
    df["MONTH"] = df["DATETIME"].dt.month.astype('int8')
    df["YEAR"] = df["DATETIME"].dt.year.astype('int16')
    df["ISO_YEAR"] = df["DATETIME"].dt.isocalendar().year.astype('int16')
    df["QUARTER"] = df["DATETIME"].dt.quarter.astype('int8')
    df["DAYOFYEAR"] = df["DATETIME"].dt.dayofyear.astype('int8')
    df["DAYOFMONTH"] = df["DATETIME"].dt.day.astype('int8')
    df["ISO_WEEKOFYEAR"] = df["DATETIME"].dt.isocalendar().week.astype('int8')
    df["SEASON"] = df["DATETIME"].dt.month%12 // 3 + 1 # Season 1 is Summer
    df["SEASON"] = df["SEASON"].astype('int8')
    return df

df_tempTd_pd = create_new_features(df_tempTd_pd)
df_tempTd_pd.info()
```


```{python Inspecting data #2}
df_tempTd_pd.head()
```


```{python Saving transformed data to new file}
# Saving data to new file as converting to R changes a few of the datetimes
#df_tempTd_pd.to_csv('../data/processed/data_used_for_EDA.csv', index=False)
```

# Visualisations 

```{r Loading new tibble, include=FALSE}
tempTd_fourStates2 <- read_csv("../data/processed/data_used_for_EDA.csv", 
                               col_types = cols(
                                  DATETIME = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
                                  REGIONID = col_factor(),
                                  HOUR = col_factor(), 
                                  ISO_DAYOFWEEK = col_factor(), 
                                  MONTH = col_factor(), 
                                  YEAR = col_factor(), 
                                  ISO_YEAR = col_factor(), 
                                  QUARTER = col_factor(), 
                                  DAYOFYEAR = col_factor(), 
                                  DAYOFMONTH = col_factor(), 
                                  ISO_WEEKOFYEAR = col_factor(), 
                                  SEASON = col_factor()
                                  )
                              )

df_mvs <- read_csv('../data/processed/data_used_for_EDA_missing_values.csv')
``` 


```{r Inspecting new tibble}
tempTd_fourStates2
df_mvs
```

# Visualisation 1 - Pre/Post-Resample

```{r}
df_mvs %>%
  mutate(variable = as.factor(case_when(variable == 'num_missing_pre' ~ 'pre-resample',
                                        variable == 'num_missing_post' ~ 'post-resample'))) %>%
  ggplot(aes(x=fct_reorder(REGIONID, desc(value)), y=value, 
             fill=fct_reorder(variable, desc(value)))) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=value), vjust = -0.2, position = position_dodge(0.9)) +
  labs(x='Region',
       y='Number of Missing Values',
       title='Missing Temperature Observations Pre and Post Resample') +
  guides(fill=guide_legend("Pre/Post Resample")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5)) +
  scale_fill_manual(values = c("pre-resample"="#00BFC4",
                               "post-resample"="#F8766D"))

#ggsave('GroupReportTemplate/EDA/fourStates_tempDataset_prePostResample.png')
```


# Visualisation 2 - Temperature vs Electricity Consumption for each Region

```{r Scatterplot of Temperature vs Demand}
tempTd_fourStates2 %>% 
  ggplot(aes(x=TEMPERATURE, y=TOTALDEMAND, color=REGIONID)) +
  geom_point() +
  facet_wrap(~REGIONID, nrow=2, ncol=2) +
  labs(x='Temperature (°C)',
       y='Electricity Consumption (MW)',
       title='Temperature vs Electricity Consumption per Region') +
  guides(color=guide_legend("Region")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_td_vs_temp_scatterplot.png')
```


# Visualisation 3 - Heatmap of Bin Counts for each Region

```{r}
tempTd_fourStates2 %>% 
  ggplot(aes(x=TEMPERATURE, y=TOTALDEMAND)) +
  geom_bin_2d(bins=30) +
  scale_fill_viridis_c() +
  facet_wrap(~REGIONID, nrow=2, ncol=2) +
  labs(x='Temperature (°C)',
       y='Electricity Consumption (MW)',
       title='Heatmap of Bin Counts for each Region') +
  guides(guide_legend(title="Count")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_td_vs_temp_HeatmapBinCounts.png')
```


# Visualisation 4 - Temperature vs Electricity Cosumption per Season for each Region 

```{r Scatterplot of Temperature vs Demand by Region and Season}
tempTd_fourStates2 %>% 
  mutate(SEASON = case_when(SEASON == 1 ~ 'Summer',
                            SEASON == 2 ~ 'Autumn',
                            SEASON == 3 ~ 'Winter',
                            SEASON == 4 ~ 'Spring')) %>% 
  mutate(SEASON = fct_relevel(SEASON, c('Summer', 'Autumn', 'Winter', 'Spring'))) %>% 
  ggplot(aes(x=TEMPERATURE, y=TOTALDEMAND, color=REGIONID)) +
  geom_point() +
  facet_grid(rows=vars(REGIONID), cols=vars(SEASON)) +
  labs(x='Temperature (°C)',
       y='Electricity Consumption (MW)',
       title='Temperature vs Electricity Consumption per Season and Region') +
  guides(color=guide_legend("Region")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_td_vs_temp_SeasonScatterplot.png')
```


# Visualisation 5 - Boxplot of Electricity Consumption per Region

```{r Boxplot of Total Demand per Region}
tempTd_fourStates2 %>% 
  ggplot(aes(x=REGIONID, y=TOTALDEMAND, fill=REGIONID)) +
  stat_boxplot(geom='errorbar', width=0.25) +
  geom_boxplot() +
  labs(x='Region',
       y='Electricity Consumption (MW)',
       title='Boxplot of Electricity Consumption per Region') +
  guides(fill=guide_legend("Region")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_td_Boxplot.png')
```


# Visualisation 6 - Boxplot of Temperature per Region

```{r Boxplot of Temperatures per Region}
tempTd_fourStates2 %>% 
  ggplot(aes(x=REGIONID, y=TEMPERATURE, fill=REGIONID)) +
  stat_boxplot(geom='errorbar', width=0.25) +
  geom_boxplot() +
  labs(x='Region',
       y='Temperature (°C)',
       title='Boxplot of Temperature per Region') +
  guides(fill=guide_legend("Region")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_temp_Boxplot.png')
```


# Visualisation 7 - Hourly Electricity Consumption per Region

```{r Bar Chart of Average Total Demand per Hour per region}
tempTd_fourStates2 %>% 
  group_by(REGIONID, HOUR) %>% 
  summarise(AVERAGE=mean(TOTALDEMAND)) %>% 
  ggplot(aes(x=HOUR, y=AVERAGE, fill=REGIONID)) +
  geom_col() +
  facet_grid(rows=vars(REGIONID), scales='free') +
  labs(title='Hourly Electricity Consumption per Region',
       x='Hour',
       y='Electricity Consumption (MW)') +
  guides(fill=guide_legend("Region")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_hour_BarChart.png')
```


# Visualisation 8 - Seasonal average Electricity Consumption per Region

```{r Bar Chart of DOW vs Average Elec Consump by Region and Season}
tempTd_fourStates2 %>% 
  mutate_at('ISO_DAYOFWEEK', ~fct_relevel(., sort(levels(.)))) %>%
  group_by(REGIONID, SEASON, ISO_DAYOFWEEK) %>% 
  summarise(AVERAGE=mean(TOTALDEMAND)) %>% 
  mutate(SEASON = as.factor(case_when(SEASON == 1 ~ 'Summer',
                                      SEASON == 2 ~ 'Autumn',
                                      SEASON == 3 ~ 'Winter',
                                      SEASON == 4 ~ 'Spring'))) %>% 
  ggplot(aes(x=ISO_DAYOFWEEK, y=AVERAGE, fill=REGIONID)) +
  geom_col() +
  facet_grid(cols=vars(SEASON), rows=vars(REGIONID), scales='free') +
  labs(title='Seasonal Average Electricity Consumption per Region',
       x='Day of Week',
       y='Average Electricity Consumption (MW)') +
  guides(fill=guide_legend("Region")) + # Sets legend title
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_seasonElecConsumpDOW_barPlot.png')
```


# Visualisation 9 - Density Plots of Electricity Consumption for each Region

```{r}
tempTd_fourStates2 %>% 
  rename(Region = REGIONID) %>% 
  ggplot(aes(x=TOTALDEMAND, color=Region, fill=Region)) +
  geom_density(alpha=0.4) +
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", binwidth=200)+
  ggtitle('Density Plot of Electricity Consumption per Region') +
  labs(x='Electricity Consumption (MW)',
       y='Density') +
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/fourStates_ElecConsump_densityPlot.png')
```


# Visualisation 10 - Christmas Day Temperature - Years 2016-2020

```{r}
# Line plot of temperatures for every Christmas
tempTd_fourStates2 %>%
  mutate(HOUR_MINUTE = as.factor(format(tempTd_fourStates2$DATETIME, '%H:%M'))) %>% 
  filter(as.numeric(MONTH) == 12,
         as.numeric(DAYOFMONTH) == 25,
         YEAR %in% c(2016, 2017, 2018, 2019, 2020)) %>% 
  ggplot(aes(x=HOUR_MINUTE, y=TEMPERATURE, group=YEAR, color=YEAR)) +
  geom_line() +
  facet_grid(rows=vars(REGIONID), scales='free') +
  theme(plot.title = element_text(hjust=0.5)) +
  guides(color=guide_legend("Year")) +
  scale_x_discrete(breaks = c("00:00", "01:00", "02:00", "03:00", "04:00",
                              "05:00", "06:00", "07:00", "08:00", "09:00",
                              "10:00", "11:00", "12:00", "13:00", "14:00",
                              "15:00", "16:00", "17:00", "18:00", "19:00",
                              "20:00", "21:00", "22:00", "23:00"),
                   labels = c("0", "1", "2", "3", "4", "5",
                              "6", "7", "8", "9", "10", "11",
                              "12", "13", "14", "15", "16", "17",
                              "18", "19", "20", "21", "22", "23")) +
  scale_y_continuous(breaks = c(15, 20, 25, 30, 35)) +
  labs(x='Time (Hours)',
       y='Temperature (°C)',
       title='Christmas Temperature - Year 2016-2020')

#ggsave('GroupReportTemplate/EDA/fourStates_xmasTemp_LinePLot.png')
```


# Visualisation 11 - Christmas Day Electricity Consumption - Years 2016-2020

```{r}
# Line plot of temperatures for every Christmas
tempTd_fourStates2 %>%
  mutate(HOUR_MINUTE = as.factor(format(tempTd_fourStates2$DATETIME, '%H:%M'))) %>% 
  filter(as.numeric(MONTH) == 12,
         as.numeric(DAYOFMONTH) == 25,
         YEAR %in% c(2016, 2017, 2018, 2019, 2020)) %>% 
  ggplot(aes(x=HOUR_MINUTE, y=TOTALDEMAND, group=YEAR, color=YEAR)) +
  geom_line() +
  facet_grid(rows=vars(REGIONID), scales='free') +
  theme(plot.title = element_text(hjust=0.5)) +
  guides(color=guide_legend("Year")) +
  scale_x_discrete(breaks = c("00:00", "01:00", "02:00", "03:00", "04:00",
                              "05:00", "06:00", "07:00", "08:00", "09:00",
                              "10:00", "11:00", "12:00", "13:00", "14:00",
                              "15:00", "16:00", "17:00", "18:00", "19:00",
                              "20:00", "21:00", "22:00", "23:00"),
                   labels = c("0", "1", "2", "3", "4", "5",
                              "6", "7", "8", "9", "10", "11",
                              "12", "13", "14", "15", "16", "17",
                              "18", "19", "20", "21", "22", "23")) +
  #scale_y_continuous(breaks = c(15, 20, 25, 30, 35)) +
  labs(x='Time (Hours)',
       y='Electricity Consumption (MW)',
       title='Christmas Electricity Consumption - Year 2016-2020')

#ggsave('GroupReportTemplate/EDA/fourStates_xmasElecCons_LinePLot.png')
```


# Visualisation 12 - Heatmap of Correlations for QLD

```{r Correlation plot between features from one Region (QLD)}
heatmap_plot <- function(df, region) {
  correlations <- df %>% 
    filter(REGIONID == {{region}}) %>% 
    select(-c('DATETIME', 'REGIONID')) %>%
    mutate(across(everything(.), as.numeric)) %>% 
    cor(.)
  
  correlations[lower.tri(correlations)] <- NA
  
  correlations[,nrow(correlations):1] %>% 
    melt(., na.rm=TRUE) %>% 
    mutate(across(value, ~round(., 2))) %>% 
    ggplot(aes(x=Var1, y=Var2, fill=value)) +
    geom_tile(color='white') +
    geom_text(aes(label=value), color='white', size=3) +
    theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(hjust=0.5)) +
    scale_fill_continuous(type="gradient", limit=c(-1,1)) +
    labs(title=glue('Heatmap of Correlations for {region}'))
}

heatmap_plot(tempTd_fourStates2, 'QLD')
#ggsave('GroupReportTemplate/EDA/QLD_correlation_plot.png')
#heatmap_plot(tempTd_fourStates2, 'VIC')
#heatmap_plot(tempTd_fourStates2, 'NSW')
#heatmap_plot(tempTd_fourStates2, 'SA')
```


# Visualisation 13 - 3-Fold Nested CV

```{r}
tibble(
  CV_Fold = c('1', '1', '1', '1', '1',
              '2', '2', '2', '2', '2',
              '3', '3', '3', '3', '3'),
  Set = fct_reorder(c("train", "train", "validation", "test", "no data",
                      "train", "train", "validation", "test", "no data",
                      "train", "train", "validation", "test", "no data"), 
                    c(1903, 1096, 731, 366, 0,
                      1903, 1096, 731, 366, 0,
                      1903, 1096, 731, 366, 0)),
  DaysInYear = c(0, 366, 365, 365, 807,
                 0, 731, 365, 365, 442,
                 0, 1096, 365, 442, 0)
) %>% 
  ggplot(aes(x=CV_Fold, y=DaysInYear, fill=Set)) +
  geom_col(width=0.6) +
  coord_flip() +
  scale_y_continuous(breaks = c(0, 366, 731, 1096, 1461, 1903),
                     labels = c("2016-01-01", "2017-01-01", "2018-01-01",
                                "2019-01-01", "2020-01-01", "2021-03-18")) +
  scale_fill_manual(values = c("train"="#F8766D",
                               "validation"="#7CAE00",
                               "test"="#00BFC4",
                               "no data"="grey")) +
  scale_x_discrete(limits=rev) +
  labs(x='CV Fold',
       y='Date',
       title='3-Fold Nested Cross Validation') +
  theme(plot.title = element_text(hjust=0.5))

#ggsave('GroupReportTemplate/EDA/3fold_nested_cv.png')
```

